### 2024-09-18 Lecture_2_Beta_Distribution.Rmd
Beta distribution example for Biostatics class lecture 2.

---
title: "Lecture 2 Distribution"
author: "Marqus Odisho"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment:
### What distribution are your data likely from?

My data is likely from a beta distribution.

### Why do you think this?

My reasoning for this is because when looking admixture, alpha and beta can be used as parameters for two different parental alleles, in my case, from different species.

### Using the distribution that you've chosen, change the different parameters. What happens when you change the parameters for the distribution? Describe how the shape changes.

As the parameters are reduced to a number below 1, the distribution begins to take on a u-shape with that shape being most prominent when both values are small. 
If one of the parameters is larger than the other, the distribution skews in the direction of the larger number.
As the parameters are increased to a number larger than 1, the distribution takes on a more n-shaped curve similar to a normal distribution. As the parameters are increased, the curve becomes wider.

### Can you simulate data that looks like what you expect your data looks like? What are the parameters for that?

```{r Beta}
### Beta Distribution
hist(rbeta(1000, 0.5, 0.5))
plot(density(rbeta(1000, 0.3, 0.3)))
```

### 2024-10-29 Paper_Summary.Rmd
Start of Paper Summary Assignment. Nothing typed up yet.

---
title: "Paper Summary"
author: "Marqus Odisho"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2024-10-29 Paper_Summary.Rmd
Created Figure 1a that looks similar to the Agrawal paper.

@@
 title: "Paper Summary"
-author: "Marqus Odisho"
+author: "Marqus Odisho & Sargon"
 date: "`r Sys.Date()`"
@@
 knitr::opts_chunk$set(echo = TRUE)
+#install.packages("ggplot2")
+#install.packages("lme4")
+#install.packages("lmerTest")
+#install.packages("MuMIn")
+#install.packages("MASS")
+#install.packages("dplyr")
+#install.packages("tidyr")
+library(ggplot2)
+library(lme4)
+library(lmerTest)
+library(MuMIn)
+library(MASS)
+library(dplyr)
+library(tidyr)
 ```
 
-## R Markdown
+```{r assay 1 simulation}
+###Trait of Interest: Creating normally distributed female fly populations with average value corresponding to their average number of sexual interactions.
+Four populations corresponding to the 4 different combinations between female fly size and environment type.
+(small females in cages, small females in vials, large females in cages, large females in vials)
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+# Sample Data (replace with your actual data)
+data <- data.frame(
+  Environment = factor(rep(c("Vial", "Cage"), each = 2), levels = c("Vial", "Cage")),
+  Size = factor(rep(c("Small", "Large"), 2), levels = c("Small", "Large")),
+  Mean = c(4, 10, 6, 5),
+  SD = c(1, 1, 1, 1)  # Standard deviation (replace with your actual values)
+)
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+# Create the bar graph with error bars (Environment on x-axis)
+ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), 
+                width = 0.2, position = position_dodge(0.9)) +
+  labs(
+    title = "Random Graph Nonsense",
+    x = "",
+    y = "Average Number of Sexual Interactions in a 1 Minute Period",
+    fill = " " 
+  ) +
+  theme_classic() +
+  theme(
+    legend.position = "bottom",
+  ) +
+  scale_fill_manual(values = c("#B6D0E2", "#F1E3D3"))  # Pastel blue and yellow
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2024-10-29 Lecture 2 Distribution.R
Adding missing files to github.

@@ -1,4 +1,33 @@
+---
+title: "Lecture 2 Distribution"
+author: "Marqus Odisho"
+date: '`r Sys.Date()`'
+output: html_document
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+# Assignment:
+## What distribution are your data likely from?
+
+My data is likely from a beta distribution.  
+
+## Why do you think this?
+
+My reasoning for this is because when looking admixture, alpha and beta can be used as parameters for two different parental alleles, in my case, from different species.
+
+## Using the distribution that you've chosen, change the different parameters. What happens when you change the parameters for the distribution? Describe how the shape changes.
+
+As the parameters are reduced to a number below 1, the distribution begins to take on a u-shape with that shape being most prominent when both values are small. 
+If one of the parameters is larger than the other, the distribution skews in the direction of the larger number.
+As the parameters are increased to a number larger than 1, the distribution takes on a more n-shaped curve similar to a normal distribution. As the parameters are increased, the curve becomes wider.
+
+## Can you simulate data that looks like what you expect your data looks like? What are the parameters for that?
+
 ```{r Beta}
-## My chick example isn't working for me as well right now. But, something like distribution of admixed hybrids (from two populations) is well approximated by beta. So let's say we have two interbreeding populations, and we want to know what the proportion of species A is in each individual. Beta's great for that. 
-hist(rbeta(100, 5, 5))
-plot(density(rbeta(100, 5, 5)))
+### Beta Distribution
+hist(rbeta(1000, 0.5, 0.5))
+plot(density(rbeta(1000, 0.3, 0.3)))
+```

### 2024-10-29 Paper_Summary.Rmd
Created Figure 1a that looks similar to the Agrawal paper.

@@
 title: "Paper Summary"
-author: "Marqus Odisho & Sargon"
+author: "Marqus Odisho & Sargon Yousefian"
 date: "`r Sys.Date()`"
@@
   ) +
-  scale_fill_manual(values = c("#B6D0E2", "#F1E3D3"))  # Pastel blue and yellow
+  scale_fill_manual(values = c("#B6D0E2", "#F1E3D3"))  # Pastel blue and yellow for the legend
 ```

### 2024-10-30 Lecture 2 Distribution.R
Delete Lecture 2 Distribution.R

@@ -1,33 +1,0 @@
----
-title: "Lecture 2 Distribution"
-author: "Marqus Odisho"
-date: '`r Sys.Date()`'
-output: html_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-# Assignment:
-## What distribution are your data likely from?
-
-My data is likely from a beta distribution.  
-
-## Why do you think this?
-
-My reasoning for this is because when looking admixture, alpha and beta can be used as parameters for two different parental alleles, in my case, from different species.
-
-## Using the distribution that you've chosen, change the different parameters. What happens when you change the parameters for the distribution? Describe how the shape changes.
-
-As the parameters are reduced to a number below 1, the distribution begins to take on a u-shape with that shape being most prominent when both values are small. 
-If one of the parameters is larger than the other, the distribution skews in the direction of the larger number.
-As the parameters are increased to a number larger than 1, the distribution takes on a more n-shaped curve similar to a normal distribution. As the parameters are increased, the curve becomes wider.
-
-## Can you simulate data that looks like what you expect your data looks like? What are the parameters for that?
-
-```{r Beta}
-### Beta Distribution
-hist(rbeta(1000, 0.5, 0.5))
-plot(density(rbeta(1000, 0.3, 0.3)))
-```

### 2024-10-30 Paper_Summary.Rmd
Added Figure 1b. code interpretation from original paper.

@@
 
-```{r assay 1 simulation}
+```{r Assay 1: Figure 1a. Simulation & Analysis}
 ###Trait of Interest: Creating normally distributed female fly populations with average value corresponding to their average number of sexual interactions.
@@
 ```
+
+```{r Assay 1: Figure 1b. Simulation & Analysis}
+MeanFeeding_Small<-c(10,9)
+MeanFeeding_Large<-c(6.5,13.5)
+EnvironmentType<-c('Vial','Cage')
+x<-c(1,2)
+SD = c(1, 1, 1, 1)
+
+plot(MeanFeeding_Small, axes=FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim=c(4,16), pch=15, col="#B6D0E2", type='o', lwd=2, cex=1.5)
+axis(2)
+axis(1, at=seq_along(MeanFeeding_Small), labels = as.character(EnvironmentType), las=0)
+lines(c(1,2), MeanFeeding_Large, col="#F1E3D3", pch=17, type="o", lty = 2, lwd=2, cex=1.5)
+
+# Vertical error bars the good ole fashion way
+arrows(x0=x, y0=MeanFeeding_Small-SD, x1=x, y1=MeanFeeding_Small+SD, code=3, angle=90, length=0.1)
+arrows(x0=x, y0=MeanFeeding_Large-SD, x1=x, y1=MeanFeeding_Large+SD, code=3, angle=90, length=0.1)
+
+# Creating a legend that labels the female size types with different colours - corresponds to the colours in the previous graph
+legend('topleft',inset=0.05,c("Small","Large"), lty=c(1,2), col=c("#B6D0E2","#F1E3D3"), lwd=2, title="Female Size")
+
+#Turn plot outline into a box to make it more cutesy
+box()
+```

### 2024-10-30 Paper_Summary.Rmd
Added Figure 1b. code interpretation from original paper.

@@
     legend.position = "bottom",
+    panel.border = element_rect(colour = "black", fill=NA)
   ) +
@@
 
-#Turn plot outline into a box to make it more cutesy
+# Turn plot outline into a box to make it more cutesy
 box()

### 2024-10-30 Paper_Summary.Rmd

--- 2024-10-29 Paper_Summary.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-#install.packages("ggplot2")
-#install.packages("lme4")
-#install.packages("lmerTest")
-#install.packages("MuMIn")
-#install.packages("MASS")
-#install.packages("dplyr")
-#install.packages("tidyr")
+#install.packages('ggplot2')
+#install.packages('lme4')
+#install.packages('lmerTest')
+#install.packages('MuMIn')
+#install.packages('MASS')
+#install.packages('dplyr')
+#install.packages('tidyr')
 library(ggplot2)
@@
 data <- data.frame(
-  Environment = factor(rep(c("Vial", "Cage"), each = 2), levels = c("Vial", "Cage")),
-  Size = factor(rep(c("Small", "Large"), 2), levels = c("Small", "Large")),
+  Environment = factor(rep(c('Vial', 'Cage'), each = 2), levels = c('Vial', 'Cage')),
+  Size = factor(rep(c('Small', 'Large'), 2), levels = c('Small', 'Large')),
   Mean = c(4, 10, 6, 5),
@@
 # Create the bar graph with error bars (Environment on x-axis)
-ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
-  geom_bar(stat = "identity", position = "dodge") +
+ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0, 2, 4, 6, 8, 10)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), 
@@
     title = "Random Graph Nonsense",
-    x = "",
+    x = '',
     y = "Average Number of Sexual Interactions in a 1 Minute Period",
-    fill = " " 
+    fill = ' ' 
   ) +
@@
   theme(
-    legend.position = "bottom",
+    legend.position = 'bottom',
+    panel.border = element_rect(colour = 'black', fill = NA)
   ) +
-  scale_fill_manual(values = c("#B6D0E2", "#F1E3D3"))  # Pastel blue and yellow for the legend
+  scale_fill_manual(values = c('#B6D0E2', '#F1E3D3'))  # Pastel blue and yellow for the legend
 ```
+
+```{r Assay 1: Figure 1b. Simulation & Analysis}
+
+#Creating useful variables to create the plot for Figure 1b.
+MeanFeeding_Small<-c(10, 9)
+MeanFeeding_Large<-c(6.5, 13.5)
+EnvironmentType<-c('Vial', 'Cage')
+x <- c(1, 2)
+SD = c(1, 1, 1, 1)
+
+plot(MeanFeeding_Small, axes=FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim = c(4, 16), pch = 15, col = '#B6D0E2', type = 'o', lwd = 2, cex = 1.5)
+axis(2)
+axis(1, at = seq_along(MeanFeeding_Small), labels = as.character(EnvironmentType), las = 0)
+lines(c(1, 2), MeanFeeding_Large, col='#F1E3D3', pch = 17, type = 'o', lty = 2, lwd = 2, cex = 1.5)
+
+# Vertical error bars the good ole fashion way
+arrows(x0 = x, y0 = MeanFeeding_Small - SD, x1 = x, y1 = MeanFeeding_Small + SD, code = 3, angle = 90, length = 0.1)
+arrows(x0 = x, y0 = MeanFeeding_Large - SD, x1 = x, y1 = MeanFeeding_Large + SD, code = 3, angle = 90, length = 0.1)
+
+# Creating a legend that labels the female size types with different colours - corresponds to the colours in the previous graph
+legend('topleft', inset = 0.05, c('Small', 'Large'), lty = c(1, 2), col =c ('#B6D0E2','#F1E3D3'), lwd = 2, title = "Female Size")
+
+# Turn plot outline into a box to make it more cutesy
+box()
+```

### 2024-10-31 Paper_Summary.Rmd
Added Sargon's Assay 1 data analysis portion. Also modified the design of the plots to be more consistent.

@@
 ```{r Assay 1: Figure 1a. Simulation & Analysis}
-###Trait of Interest: Creating normally distributed female fly populations with average value corresponding to their average number of sexual interactions.
-Four populations corresponding to the 4 different combinations between female fly size and environment type.
-(small females in cages, small females in vials, large females in cages, large females in vials)
+# Trait of Interest:
 
@@
   labs(
-    title = "Random Graph Nonsense",
+    title = '(a)',
     x = '',
@@
   theme(
+    plot.title = element_text(hjust = 0.5),
     legend.position = 'bottom',
@@
 
-#Creating useful variables to create the plot for Figure 1b.
+# Spawning and simulating useful variables to create the plot for Figure 1b
 MeanFeeding_Small<-c(10, 9)
@@
 
-plot(MeanFeeding_Small, axes=FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim = c(4, 16), pch = 15, col = '#B6D0E2', type = 'o', lwd = 2, cex = 1.5)
+# Creating a plot that is similar to Figure 1b. in the paper
+plot(MeanFeeding_Small, axes = FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim = c(4, 16), pch = 15, col = '#B6D0E2', type = 'o', lwd = 2, cex = 1.5)
   axis(2)
@@
   lines(c(1, 2), MeanFeeding_Large, col='#F1E3D3', pch = 17, type = 'o', lty = 2, lwd = 2, cex = 1.5)
+  title(main = '(b)')
 
   # Vertical error bars the good ole fashion way
-arrows(x0 = x, y0 = MeanFeeding_Small - SD, x1 = x, y1 = MeanFeeding_Small + SD, code = 3, angle = 90, length = 0.1)
-arrows(x0 = x, y0 = MeanFeeding_Large - SD, x1 = x, y1 = MeanFeeding_Large + SD, code = 3, angle = 90, length = 0.1)
+  arrows(x0 = x, y0 = MeanFeeding_Small - SD, x1 = x, y1 = MeanFeeding_Small + SD, code = 3, angle = 90, 
+         length = 0.1)
+  arrows(x0 = x, y0 = MeanFeeding_Large - SD, x1 = x, y1 = MeanFeeding_Large + SD, code = 3, angle = 90, 
+         length = 0.1)
 
   # Creating a legend that labels the female size types with different colours - corresponds to the colours in      the previous graph
-legend('topleft', inset = 0.05, c('Small', 'Large'), lty = c(1, 2), col =c ('#B6D0E2','#F1E3D3'), lwd = 2, title = "Female Size")
+  legend('topleft', inset = 0.05, c('Small', 'Large'), lty = c(1, 2), col =c ('#B6D0E2','#F1E3D3'), 
+         lwd = 2, title = "Female Size")
 
@@
 ```
+```{r Assay 2: Figure 2. Simulation & Analysis}
+low.vial<-rnorm(35, 1, 0.1)
+high.vial<-rnorm(35, 1, 0.1)
+low.cage<-rnorm(35, 1, 0.1)
+high.cage<-rnorm(35, 1, 0.1)
+g<-sample(1:4, 140, replace=TRUE)
+error<-ifelse(g==1, 0.1, ifelse(g==2, 0.5, 1.5))
+g<-as.factor(g)
+response<-24*low.vial+17*high.vial+30*low.cage+28*high.cage+rnorm(140, mean=0, sd=0.1)
+data<-cbind.data.frame(response, low.vial, high.vial, low.cage, high.cage)
+model1<-glm(response~low.vial+high.vial+low.cage+high.cage, data=data)
+summary(model1)
+anova(model1)
+```

### 2024-11-02 Paper_Summary.Rmd
Added new test code. Note: Not fully working yet.

@@
 #install.packages('tidyr')
+#install.packages('car')
+#install.packages('pbkrtest')
+#install.packages('pscl')
 library(ggplot2)
@@
 library(tidyr)
+library(car)
+library(pbkrtest)
+library(pscl)
 ```
 
-```{r Assay 1: Figure 1a. Simulation & Analysis}
+```{r Assay 1: Figure 1a. Data Simulation & Analysis}
+
+```
+
+```{r test}
+# Simulate Trait of Interest
+rnorm(1000, 10, 1)->trait_of_interest ### change this to be a trait that you're actually interested in, with an appropriate distribution!
+
+plot(density(trait_of_interest))
+min(trait_of_interest)
+max(trait_of_interest)
+
+yes<-sample(trait_of_interest, 500, replace=FALSE, prob = ifelse(trait_of_interest>0, 0.95, 0.15)) #### play with this line! Is the test statistically significant. When is it not?
+predictor_t.test<-(trait_of_interest %in% yes)
+
+cbind.data.frame(trait_of_interest, predictor_t.test)->data
+
+mean(data[which(data$predictor_t.test==TRUE),1])
+mean(data[which(data$predictor_t.test==FALSE),1])
+t.test(trait_of_interest~predictor_t.test, data=data) ### this does a two sample t-test. What would a one sample t test be testing? How would you do that?
+
+### plots our two samples for distribution
+plot(density(data[which(data$predictor_t.test==FALSE),1]), col="red", main="Two sample t test")
+lines(density(data[which(data$predictor_t.test==TRUE),1]), ylim=c(0, 0.1), xlim=c(-20,20), main="Two Sample T test")
+
+test1<-sample(trait_of_interest, 200, replace=FALSE, prob = ifelse(trait_of_interest>7, 0.95, 0.15))
+test2<-sample(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)] >4, 0.95, 0.15))
+test3<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 == FALSE)] >0, 0.95, 0.15))
+test4<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)] >-4, 0.95, 0.15))
+test5<-trait_of_interest[which(trait_of_interest %in% test1 == FALSE| trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test3 == FALSE | trait_of_interest %in% test4 == FALSE)]
+
+plot(density(test1), ylim=c(0, 0.1), main="Anovas")
+lines(density(test2), col="red")
+lines(density(test3), col="blue")
+lines(density(test4), col="purple")
+lines(density(test5), col="yellow")
+
+anova_predictor<-data.frame(ifelse(trait_of_interest %in% test1 == TRUE, "group1", ifelse(trait_of_interest %in% test2 == TRUE, "group2", ifelse(trait_of_interest %in% test3 == TRUE, "group3", ifelse(trait_of_interest %in% test4 == TRUE, "group4", "group5")))))
+
+data2<-cbind.data.frame(data, anova_predictor)
+names(data2)<-c("trait_of_interest", "predictor_t.test", "anova_predictor")
+
+anova(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do?
+
+anova(lm(trait_of_interest~anova_predictor, data=data2)) ### what does this do? 
+summary(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do? What do you notice about the last three tests?
+
+```
+
+```{r Assay 1: Figure 1a. Plot}
 # Trait of Interest:
@@
 
-```{r Assay 1: Figure 1b. Simulation & Analysis}
+```{r Assay 1: Figure 1b. Data Simulation & Analysis}
+# RNG time! - Let's keep it consistent 
+set.seed(123)
+
+# Number of observations made - 2 females, 3 times a day for 3 days * ten replicates for each arena = 180 observations per environment type.
+N = 180
+
+# Changing the effect sizes to get the same values as the paper
+effect_arena <- 2.0       # Main effect of Arena
+effect_size <- 0.5        # Main effect of Size
+effect_interaction <- 2.0  # Interaction effect
+
+# Data simulation time! - Different means for different groups
+feeding_count <- c(
+  rnorm(N, mean = 6.5, sd = 1), # Large females in vials
+  rnorm(N, mean = 13.5 + effect_arena, sd = 1), # Large females in cages
+  rnorm(N, mean = 10 + effect_size, sd = 1), # Small females in vials
+  rnorm(N, mean = 8.5 + effect_arena + (effect_size * (1 + effect_interaction)), sd = 1) # Small females in cages
+)
+
+# Making sure that FeedingCount is a positive float (can't eat negative amounts!)
+feeding_count <- ifelse(feeding_count <= 0, feeding_count + 0.1, feeding_count)
+
+# Let's make all of this into a data frame - matrix time :(
+data <- data.frame(
+  Arena = factor(rep(c('vial', 'cage'), each = N * 2)),
+  Size = factor(rep(c('large', 'small'), times = N * 2)),
+  FeedingCount = feeding_count
+)
+
+model_gaussian <- glm(FeedingCount ~ Arena * Size, data = data, family = gaussian(link = 'log'))
+
+summary(model_gaussian)
+
+# Use lmerTest to get F-statistics and p-values
+anova_result <- anova(model_gaussian, test = "F")
+
+# Print the results
+print(anova_result)
+
+```
+
+```{r Assay 1: Figure 1b. Plot}
 
@@
 ```
-```{r Assay 2: Figure 2. Simulation & Analysis}
+
+```{r Assay 2: Figure 2. Data Simulation & Analysis}
 low.vial<-rnorm(35, 1, 0.1)

### 2024-11-05 Paper_Summary.Rmd
Added code for data simulation and analysis for Figure 1a. Removed test code and updated the code for Figure 2's data simulation and analysis (with results interpretations). Also added portions of methods summary and statistical methods summary while improving overall formatting.

@@
 ```
+## Data Simulation & Analysis
+### Assay 1:
+For the data 
+```{r Assay 1: Figure 1a. Data Simulation & Analysis}
+# Set the number of replicates from paper
+n_vials <- 10
 
-```{r Assay 1: Figure 1a. Data Simulation & Analysis}
+# Set the average differences in activity scores
+mean_diff_simple <- 6
+mean_diff_complex <- -1
+
+# Adjust the standard deviation to achieve the desired t-value and p-value
+sd_diff <- 2.5
+
+# Create an empty data frame
+activity_data <- data.frame(
+  female_type = character(0),
+  arena_type = character(0),
+  vial_cage_id = numeric(0),
+  activity_score = numeric(0)
+)
+
+# Generate data for each vial/cage (for loops are harder in R than in Python in my opinion!)
+for (i in 1:n_vials) {
+  # Simple arena
+  activity_data <- rbind(activity_data, data.frame(
+    female_type = "large",
+    arena_type = "simple",
+    vial_cage_id = i,
+    activity_score = rnorm(1, mean = 10 + mean_diff_simple/2, sd = sd_diff)
+  ))
+  activity_data <- rbind(activity_data, data.frame(
+    female_type = "small",
+    arena_type = "simple",
+    vial_cage_id = i,
+    activity_score = rnorm(1, mean = 10 - mean_diff_simple/2, sd = sd_diff)
+  ))
+
+  # Complex arena
+  activity_data <- rbind(activity_data, data.frame(
+    female_type = "large",
+    arena_type = "complex",
+    vial_cage_id = i,
+    activity_score = rnorm(1, mean = 10 + mean_diff_complex/2, sd = sd_diff)
+  ))
+  activity_data <- rbind(activity_data, data.frame(
+    female_type = "small",
+    arena_type = "complex",
+    vial_cage_id = i,
+    activity_score = rnorm(1, mean = 10 - mean_diff_complex/2, sd = sd_diff)
+  ))
+}
+
+# Calculate the difference in activity scores within each vial/cage
+activity_data <- activity_data %>%
+  group_by(arena_type, vial_cage_id) %>%
+  summarize(score_difference = activity_score[female_type == "large"] - activity_score[female_type == "small"])
+
+# Perform the two-sample t-test 
+t.test(score_difference ~ arena_type, data = activity_data)
 

### 2024-11-05 Paper_Summary.Rmd
Adding more text and improving formatting.

@@
 ---
-title: "Paper Summary"
+title: "Paper Summary Assignment"
 author: "Marqus Odisho & Sargon Yousefian"
@@
 ```
-## Data Simulation & Analysis
-### Assay 1:
-For the data 
+### Paper Choice
+
+_The physical environment mediates male harm and its effect on selection in females_. 
+Li Yun, Patrick J. Chen, Amardeep Singh, Aneil F. Agrawal, and Howard D. Rundle.
+
+### Methods Summary
+
+#### Assay 1
+
+This assay examined the impact of the physical environment on male fruit fly sexual behaviors.
+Two different environments were set up: a small vial and a larger Ziploc container, both containing food.
+Each environment had ten replicates, with 37 male and 37 female flies in each.
+The female flies were divided into two categories, small and large, based on manipulated egg density.
+For each replicate, four females, two large and two small, were marked with different fluorescent powders for identification.
+The flies were allowed to acclimate for one hour before observations began.
+Researchers recorded their movements three times a day (at 7:00, 12:00, and 17:00).
+During one-minute observation periods, the sexual activities directed toward each marked female were noted, including courtship behaviors like singing and chasing, as well as copulation attempts.
+The duration of these activities and whether the females fed during the observation were also recorded.
+Data from each arena were then aggregated for analysis.
+
+#### Assay 2
+
+The second assay examined how environmental factors influence male-induced harm to female fruit flies through a two-way factorial experiment, manipulating male exposure (low vs.
+high) and environmental complexity, simple vs.
+complex.
+The simple environment was represented by vials, while the complex environment was represented by Ziploc containers.
+In total, 35 replicates were observed, each containing 35 female flies.
+In the low-exposure condition, females were exposed to males for three hours on days 1, 4, and 6 of their mating phase.
+In the high-exposure condition, males were present continuously in the arenas.
+After the third exposure, females were assessed for survival.
+From each replicate, ten random females were selected, and the number of adults or emerging pupae from their eggs was recorded.
+Total adult fitness, encompassing both fecundity and survival, was calculated by multiplying these two values.
+
+### Statistical Methods Summary
+
+#### Assay 1
+
+For this assay, two statistical methods were used to help interpret the results.
+First, a two-sample t-test treating separate mating arenas as replicates was conducted to compare the difference in sexual activity score of the two female types (large and small) between the two arenas.
+The reason the authors chose to create a sexual activity score and compare that across mating arenas was because the observations within a certain vial or cage were not independent.
+The results indicated that the means (difference in sexual activity) of the vial group and the cage group were significantly different as indicated by the values (t18 = 5.89, p < 0.001).
+To the authors this suggested that the difference in male sexual attention towards large and small females was different in vials vs.
+in cages.
+It made sense to use the two-sample t-test to compare means between the two environments as one of the goals of this paper was to compare how the environment changed the sexual interactions of Drosophila melanogaster.
+It is also important to note that the authors took steps to avoid violating one of the assumptions of the two-sample t-test (independence) by creating a single sexual activity score for each replicate (for both small and large females) as the observations of average number of sexual encounters within the cages and vials were not independent.
+
+#### Assay 2
+
+Insert sample text.
+
+### Data Simulation & Analysis
+
+#### Assay 1
+
+For the methods and results in Assay 1, the dataset was simulated and analysized in two different ways corresponding to Figures 1a and 1b in the original paper.
+The corresponding code is seperated into two blocks, one for Figure 1a and the other for Figure 1b.
+
+#### Figure 1a: Data Simulation and Analysis
+
 ```{r Assay 1: Figure 1a. Data Simulation & Analysis}
@@
 
-```{r test}
-# Simulate Trait of Interest
-rnorm(1000, 10, 1)->trait_of_interest ### change this to be a trait that you're actually interested in, with an appropriate distribution!
-
-plot(density(trait_of_interest))
-min(trait_of_interest)
-max(trait_of_interest)
-
-yes<-sample(trait_of_interest, 500, replace=FALSE, prob = ifelse(trait_of_interest>0, 0.95, 0.15)) #### play with this line! Is the test statistically significant. When is it not?
-predictor_t.test<-(trait_of_interest %in% yes)
-
-cbind.data.frame(trait_of_interest, predictor_t.test)->data
-
-mean(data[which(data$predictor_t.test==TRUE),1])
-mean(data[which(data$predictor_t.test==FALSE),1])
-t.test(trait_of_interest~predictor_t.test, data=data) ### this does a two sample t-test. What would a one sample t test be testing? How would you do that?
-
-### plots our two samples for distribution
-plot(density(data[which(data$predictor_t.test==FALSE),1]), col="red", main="Two sample t test")
-lines(density(data[which(data$predictor_t.test==TRUE),1]), ylim=c(0, 0.1), xlim=c(-20,20), main="Two Sample T test")
-
-test1<-sample(trait_of_interest, 200, replace=FALSE, prob = ifelse(trait_of_interest>7, 0.95, 0.15))
-test2<-sample(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)] >4, 0.95, 0.15))
-test3<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 == FALSE)] >0, 0.95, 0.15))
-test4<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)] >-4, 0.95, 0.15))
-test5<-trait_of_interest[which(trait_of_interest %in% test1 == FALSE| trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test3 == FALSE | trait_of_interest %in% test4 == FALSE)]
-
-plot(density(test1), ylim=c(0, 0.1), main="Anovas")
-lines(density(test2), col="red")
-lines(density(test3), col="blue")
-lines(density(test4), col="purple")
-lines(density(test5), col="yellow")
-
-anova_predictor<-data.frame(ifelse(trait_of_interest %in% test1 == TRUE, "group1", ifelse(trait_of_interest %in% test2 == TRUE, "group2", ifelse(trait_of_interest %in% test3 == TRUE, "group3", ifelse(trait_of_interest %in% test4 == TRUE, "group4", "group5")))))
-
-data2<-cbind.data.frame(data, anova_predictor)
-names(data2)<-c("trait_of_interest", "predictor_t.test", "anova_predictor")
-
-anova(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do?
-
-anova(lm(trait_of_interest~anova_predictor, data=data2)) ### what does this do? 
-summary(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do? What do you notice about the last three tests?
-
-```
+#### Figure 1a. Plot of Results
 
@@
 ```
+
+#### Interpretation of Results in Figure 1a
+
+Sample Text. Please fill in.
+
+#### Figure 1b: Data Simulation and Analysis
 
@@
 
+#### Figure 1b: Plot of Results
+
 ```{r Assay 1: Figure 1b. Plot}
@@
 
+#### Interpretation of Results in Figure 1b
+
+Sample text. Please fill in.
+
+#### Assay 2
+
+#### Figure 2: Data Simulation & Analysis
+
 ```{r Assay 2: Figure 2. Data Simulation & Analysis}
-low.vial<-rnorm(35, 1, 0.1)
-high.vial<-rnorm(35, 1, 0.1)
-low.cage<-rnorm(35, 1, 0.1)
-high.cage<-rnorm(35, 1, 0.1)
-g<-sample(1:4, 140, replace=TRUE)
-error<-ifelse(g==1, 0.1, ifelse(g==2, 0.5, 1.5))
-g<-as.factor(g)
-response<-24*low.vial+17*high.vial+30*low.cage+28*high.cage+rnorm(140, mean=0, sd=0.1)
-data<-cbind.data.frame(response, low.vial, high.vial, low.cage, high.cage)
-model1<-glm(response~low.vial+high.vial+low.cage+high.cage, data=data)
-summary(model1)
-anova(model1)
+# Simulated data
+low.vial <- rnorm(35, 1, 0.1)
+high.vial <- rnorm(35, 1, 0.1)
+low.cage <- rnorm(35, 1, 0.1)
+high.cage <- rnorm(35, 1, 0.1)
+
+# Combine into a single response variable
+response <- c(
+  24 * low.vial + rnorm(35, mean = 0, sd = 0.1),  # Low Vial
+  17 * high.vial + rnorm(35, mean = 0, sd = 0.1), # High Vial
+  30 * low.cage + rnorm(35, mean = 0, sd = 0.1),  # Low Cage
+  28 * high.cage + rnorm(35, mean = 0, sd = 0.1)   # High Cage
+)
+
+# Divide data points in groups (low vs high) and type (vial vs cage)
+data <- data.frame(
+  response = response,
+  group = factor(rep(c("Low", "High", "Low", "High"), each = 35)),
+  type = factor(rep(c("Vial", "Vial", "Cage", "Cage"), times = 35))
+  )
+
+# Adjust to ensure proper pairing of group and type
+data$type <- factor(rep(c("Vial", "Vial", "Cage", "Cage"), each = 35))
+
+# Fit the two-way ANOVA model
+anova_model <- aov(response ~ group + type + group:type, data = data)
+
+# Summary of the ANOVA
+summary(anova_model)
+
 ```
+
+#### Figure 2: Plot of Results
+
+Insert plot here.
+
+#### Interpretation of Results in Figure 2
+
+From the ANOVA generated by the generalized linear model, both the P value comparing high versus low exposure and the vial versus the cage arena were lower than 0.05.
+This indicates that the there are significant differences between each treatment.
+Specifically, the statistical analysis reveals that a higher exposure to males leads to a reduced adult fitness.
+Furthermore, it also reveals that in the cage arena the females had a higher adult fitness than in the vial.

### 2024-11-05 Paper_Summary.Rmd
Starting registered report

@@
 
-# Generate data for each vial/cage (for loops are harder in R than in Python in my opinion!)
+# Generate data for each vial/cage (For loops are harder in R than in Python in my opinion!)
 for (i in 1:n_vials) {

### 2024-11-05 Registered Report.Rmd
Starting registered report

---
title: "Registered Report"
author: "Marqus Odisho"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('ggplot2')
#install.packages('lme4')
#install.packages('lmerTest')
#install.packages('MuMIn')
#install.packages('MASS')
#install.packages('dplyr')
#install.packages('tidyr')
#install.packages('car')
#install.packages('pbkrtest')
#install.packages('pscl')
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(MASS)
library(dplyr)
library(tidyr)
library(car)
library(pbkrtest)
library(pscl)
```

```{r Data Simulation}
q_maternal <- rbeta(1000, 0.5, 0.5)
q_paternal <- rbeta(1000, 0.5, 0.5)
plot(density(q_maternal))

# Add correlation test at one point
```

### 2024-11-07 Registered Report.Rmd
Starting registered report
Minor updates for paper summary. For registered report, added scatterplots with q score matching that represent different "types" of assortative mating.

@@
 
+```{r test}
+# Set a seed for reproducibility
+set.seed(123)
+
+# Create two datasets, `q_maternal` and `q_paternal`, each with 1000 data points following a beta distribution
+q_maternal <- rbeta(1000, 1, 1)
+q_paternal <- rbeta(1000, 1, 1)
+
+# Create a new dataset, `sampled_q_maternal`, by sampling 500 q scores without replacement from the `q_maternal` dataset
+sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+
+# Create a new variable, `sampled_q_paternal`, by sampling 500 q scores from the `q_paternal` dataset
+# Use the `ifelse()` function to set the probability of selecting a q score from `q_paternal` based on the corresponding q score in `sampled_q_maternal`
+sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
+  if (x >= 0.5) {
+    sample(q_paternal[q_paternal >= 0.5], 1, prob = rep(0.9 / length(q_paternal[q_paternal >= 0.5]), length(q_paternal[q_paternal >= 0.5])))
+  } else {
+    sample(q_paternal[q_paternal < 0.5], 1, prob = rep(0.1 / length(q_paternal[q_paternal < 0.5]), length(q_paternal[q_paternal < 0.5])))
+  }
+})
+
+# Create a new dataframe, `q_scores_df`, by combining `sampled_q_maternal` and `sampled_q_paternal`
+q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
+
+# Plot a scatterplot of `sampled_q_paternal` versus `sampled_q_maternal` to visualize the relationship between the two variables
+plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+```
+
+```{r test 2}
+# Set a seed for reproducibility
+set.seed(123)
+
+# Create two datasets, `q_maternal` and `q_paternal`, each with 1000 data points following a beta distribution
+q_maternal <- rbeta(1000, 1, 1)
+q_paternal <- rbeta(1000, 1, 1)
+
+# Create a new dataset, `sampled_q_maternal`, by sampling 500 q scores without replacement from the `q_maternal` dataset
+sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+
+# Create a new variable, `sampled_q_paternal`, by sampling 500 q scores from the `q_paternal` dataset
+# Use the `ifelse()` function to set the probability of selecting a q score from `q_paternal` based on the corresponding q score in `sampled_q_maternal`
+sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
+  if (x <= 0.5) {
+    # If the maternal q score is less than or equal to 0.5, sample from the lower half of the paternal q scores with higher probability
+    sample(q_paternal[q_paternal <= 0.5], 1, prob = ifelse(q_paternal[q_paternal <= 0.5] <= 0.25, 0.9, 0.1))
+  } else {
+    # If the maternal q score is greater than 0.5, sample from the upper half of the paternal q scores with higher probability
+    sample(q_paternal[q_paternal > 0.5], 1, prob = ifelse(q_paternal[q_paternal > 0.5] >= 0.75, 0.9, 0.1))
+  }
+})
+
+# Create a new dataframe, `q_scores_df`, by combining `sampled_q_maternal` and `sampled_q_paternal`
+q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
+
+# Plot a scatterplot of `sampled_q_paternal` versus `sampled_q_maternal` to visualize the relationship between the two variables
+plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+```
 ```{r Data Simulation}
-q_maternal <- rbeta(1000, 0.5, 0.5)
-q_paternal <- rbeta(1000, 0.5, 0.5)
-plot(density(q_maternal))
+# Simulate q score data for maternal and paternal species in a beta distribution with a random mating pattern to draw samples from. 
+q_maternal <- rbeta(1000, 1, 1)
+q_paternal <- rbeta(1000, 1, 1)
+
+# Visualize random mating pattern following a beta distribution
+hist(q_paternal)
+hist(q_maternal)
+
+# Sample and group q scores to indicate predicted mating patterns 
+maternal_random <- sample(q_maternal, 500, replace = FALSE)
 

### 2024-11-12 Registered Report.Rmd
Added descriptions of the code.

--- 2024-11-05 Registered Report.Rmd
@@
 
-```{r Data Simulation}
-q_maternal <- rbeta(1000, 0.5, 0.5)
-q_paternal <- rbeta(1000, 0.5, 0.5)
-plot(density(q_maternal))
+### Description
 
-# Add correlation test at one point
+Code looks to simulate assortative mating between two species to form hybrids.
+This assortative mating will be based on q scores which can range from 0 (which indicates species 1) to 1 (which indicates species 2) and intermediate values which indicate hybrids of the two.
+The code will simulate two types of assortative mating, the strict definition which means that individuals will want to mate with someone with an identical q score and the extreme phenotype definition which is where individuals will want to mate with someone with either a higher or lower q score than them.
+The code will simulate individual maternal and paternal datasets and then pair them up based on either of the two assortative mating definitions previously mentioned.
+The aformentioned pairing will be visualized in a scatterplot to show the mating patterns.
+Then hybrids will be generated based on those previous mating patterns that will be a blend of the q scores of both parents.
+The q scores of the hybrids will be plotted to show how different assortative mating patterns influence hybrid zone outcomes.
+
+```{r Assortative Mating - Strict Definition}
+# Set a seed for reproducibility
+set.seed(123)
+
+# Create two datasets, q_maternal and q_paternal, each with 1000 data points following a beta distribution where the most common q scores are 1 or 0 but there are hybrids
+q_maternal <- rbeta(1000, 0.3, 0.3)
+q_paternal <- rbeta(1000, 0.3, 0.3)
+
+# Display simulated dataset distribution before sampling
+hist(q_paternal)
+hist(q_maternal)
+
+# Create a new dataset, sampled_q_maternal, by sampling 500 q scores without replacement from the q_maternal dataset
+sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+
+# Create a new variable, sampled_q_paternal, by sampling 500 q scores from the q_paternal dataset
+# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal
+sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
+  if (x >= 0.5) {
+    sample(q_paternal[q_paternal >= 0.5], 1, prob = rep(0.9 / length(q_paternal[q_paternal >= 0.5]), length(q_paternal[q_paternal >= 0.5])))
+  } else {
+    sample(q_paternal[q_paternal < 0.5], 1, prob = rep(0.1 / length(q_paternal[q_paternal < 0.5]), length(q_paternal[q_paternal < 0.5])))
+  }
+})
+
+# Create a new dataframe, `q_scores_df`, by combining sampled_q_maternal and sampled_q_paternal
+q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
+
+# Plot a scatterplot of sampled_q_paternal versus sampled_q_maternal to visualize the relationship between the two variables where mating is based on having a similar q score
+plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
 ```
+
+```{r Assortative Mating - Extreme Phenotype Preference}
+# Set a seed for reproducibility
+set.seed(123)
+
+# Create two datasets, `q_maternal` and `q_paternal`, each with 1000 data points following a beta distribution
+q_maternal <- rbeta(1000, 0.3, 0.3)
+q_paternal <- rbeta(1000, 0.3, 0.3)
+
+# Display simulated dataset distribution before sampling
+hist(q_paternal)
+hist(q_maternal)
+
+# Create a new dataset, sampled_q_maternal, by sampling 500 q scores without replacement from the q_maternal dataset
+sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+
+# Create a new variable, sampled_q_paternal, by sampling 500 q scores from the q_paternal dataset
+# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal
+sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
+  if (x <= 0.5) {
+    # If the maternal q score is less than or equal to 0.5, sample from the lower half of the paternal q scores with higher probability
+    sample(q_paternal[q_paternal <= 0.5], 1, prob = ifelse(q_paternal[q_paternal <= 0.5] <= 0.25, 0.9, 0.1))
+  } else {
+    # If the maternal q score is greater than 0.5, sample from the upper half of the paternal q scores with higher probability
+    sample(q_paternal[q_paternal > 0.5], 1, prob = ifelse(q_paternal[q_paternal > 0.5] >= 0.75, 0.9, 0.1))
+  }
+})
+
+# Create a new dataframe, q_scores_df, by combining sampled_q_maternal and sampled_q_paternal
+q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
+
+# Plot a scatterplot of sampled_q_paternal versus sampled_q_maternal to visualize the relationship between the two variables where they are paired based on extreme q score preference
+plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+```

### 2024-11-12 Registered Report.Rmd
Added density plots showing offspring q scores based on different assortative mating patterns of the parents.

@@
 The aformentioned pairing will be visualized in a scatterplot to show the mating patterns.
-Then hybrids will be generated based on those previous mating patterns that will be a blend of the q scores of both parents.
-The q scores of the hybrids will be plotted to show how different assortative mating patterns influence hybrid zone outcomes.
+Then, offspring will be generated based on those previous mating patterns that will be a blend of the q scores of both parents.
+The q scores of the offspring will be plotted to show how different assortative mating patterns influence hybrid zone outcomes.
 
+### Assortative Mating - Strict Definition
 ```{r Assortative Mating - Strict Definition}
@@
 plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+
+q_scores_df$averaged_q_scores <- (q_scores_df$sampled_q_maternal + q_scores_df$sampled_q_paternal) / 2
+# Estimate the density of averaged_q_scores
+# density_averaged_q_scores <- density(q_scores_df$averaged_q_scores)
+
+# Plot the density plot which shows offspring q scores based on extreme phenotype assortative mating
+plot(density(q_scores_df$averaged_q_scores), 
+     main = "Density Plot of Offspring q Scores",
+     xlab = "q score",
+     ylab = "Density")
+
+# Add a rug plot to the bottom of the density plot to show the individual data points that make up the distribution
+rug(q_scores_df$averaged_q_scores)
 ```
 
+### Assortative Mating - Extreme Phenotype Preference
 ```{r Assortative Mating - Extreme Phenotype Preference}
@@
 plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+
+q_scores_df$averaged_q_scores <- (q_scores_df$sampled_q_maternal + q_scores_df$sampled_q_paternal) / 2
+# Estimate the density of averaged_q_scores
+# density_averaged_q_scores <- density(q_scores_df$averaged_q_scores)
+
+# Plot the density plot which shows offspring q scores based on extreme phenotype assortative mating
+plot(density(q_scores_df$averaged_q_scores), 
+     main = "Density Plot of Offspring q Scores",
+     xlab = "q score",
+     ylab = "Density")
+
+# Add a rug plot to the bottom of the density plot to show the individual data points that make up the distribution
+rug(q_scores_df$averaged_q_scores)
 ```

### 2024-11-12 Registered Report.Rmd
Added Spearman correlation tests.

@@
 rug(q_scores_df$averaged_q_scores)
+
+# Perform Spearman rank correlation test
+cor_test_result <- cor.test(q_scores_df$sampled_q_maternal, 
+                           q_scores_df$sampled_q_paternal, 
+                           method = "spearman")
+
+# Print the correlation test result
+print(cor_test_result)
 ```

### 2024-11-12 Registered Report.Rmd
Added Spearman correlation tests.
Improved formatting and added more annotations.

@@
 
-Code looks to simulate assortative mating between two species to form hybrids.
+Code looks to simulate assortative mating between two species to potentially form hybrid offspring.
 This assortative mating will be based on q scores which can range from 0 (which indicates species 1) to 1 (which indicates species 2) and intermediate values which indicate hybrids of the two.
@@
 The q scores of the offspring will be plotted to show how different assortative mating patterns influence hybrid zone outcomes.
+A correlation test will be performed on the data to understand the relationship between the q scores.
 
@@
 ```{r Assortative Mating - Strict Definition}
-# Set a seed for reproducibility
-set.seed(123)
-
 # Create two datasets, q_maternal and q_paternal, each with 1000 data points following a beta distribution where the most common q scores are 1 or 0 but there are hybrids
@@
 
-# Display simulated dataset distribution before sampling
-hist(q_paternal)
-hist(q_maternal)
+# Display simulated dataset distributions before sampling
+hist(q_paternal,
+     main = "Paternal q scores",
+     xlab = "q scores"
+     )
+
+hist(q_maternal,
+     main = "Maternal q scores",
+     xlab = "q scores"
+     )
 
@@
 # Create a new variable, sampled_q_paternal, by sampling 500 q scores from the q_paternal dataset
-# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal
+# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal, in this case the function x represents sampled maternal q scores that are then being matched with sampled paternal q scores based on preset probabilities.
 sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
@@
 
-# Create a new dataframe, `q_scores_df`, by combining sampled_q_maternal and sampled_q_paternal
+# Create a new dataframe, q_scores_df, by combining sampled_q_maternal and sampled_q_paternal
 q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
@@
 
+# To plot offspring q scores, take the q scores of the parents from the previous data frame and average them 
 q_scores_df$averaged_q_scores <- (q_scores_df$sampled_q_maternal + q_scores_df$sampled_q_paternal) / 2
-# Estimate the density of averaged_q_scores
-# density_averaged_q_scores <- density(q_scores_df$averaged_q_scores)
+
+# Plot the density plot which shows offspring q scores based on the strict definition of assortative mating
+plot(density(q_scores_df$averaged_q_scores), 
+     main = "Density Plot of Offspring q scores",
+     xlab = "q score",
+     ylab = "Density")
+
+# Add a rug plot to the bottom of the density plot to show the individual data points that make up the distribution
+rug(q_scores_df$averaged_q_scores)
+
+# Perform Spearman rank correlation test between maternal and paternal q scores
+cor_test_result <- cor.test(q_scores_df$sampled_q_maternal, 
+                           q_scores_df$sampled_q_paternal, 
+                           method = "spearman")
+
+# Print the correlation test result
+print(cor_test_result)
+```
+
+### Assortative Mating - Extreme Phenotype Preference
+```{r Assortative Mating - Extreme Phenotype Preference}
+# Create two datasets, q_maternal and q_paternal, each with 1000 data points following a beta distribution
+q_maternal <- rbeta(1000, 0.3, 0.3)
+q_paternal <- rbeta(1000, 0.3, 0.3)
+
+# Display simulated dataset distribution before sampling
+hist(q_paternal,
+     main = "Paternal q scores",
+     xlab = "q scores"
+     )
+
+hist(q_maternal,
+     main = "Maternal q scores",
+     xlab = "q scores"
+     )
+
+# Create a new dataset, sampled_q_maternal, by sampling 500 q scores without replacement from the q_maternal dataset
+sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+
+# Create a new variable, sampled_q_paternal, by sampling 500 q scores from the q_paternal dataset
+# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal, in this case the function x represents sampled maternal q scores that are then being matched with sampled paternal q scores based on preset probabilities.
+sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
+  if (x <= 0.5) {
+    # If the maternal q score is less than or equal to 0.5, sample from the lower half of the paternal q scores with higher probability with a nested if statement where if maternal q score is less than 0.25 it has a probability of 0.9 and if it is between 0.5 and 0.25 it has a probability of 0.1
+    sample(q_paternal[q_paternal <= 0.5], 1, prob = ifelse(q_paternal[q_paternal <= 0.5] <= 0.25, 0.9, 0.1))
+  } else {
+    # If the maternal q score is greater than 0.5, sample from the upper half of the paternal q scores with higher probability with a nested if statement where if maternal q score is greater than 0.75 it has a selection probability of 0.9 and if it is between 0.5 and 0.75 it has a probability of 0.1
+    sample(q_paternal[q_paternal > 0.5], 1, prob = ifelse(q_paternal[q_paternal > 0.5] >= 0.75, 0.9, 0.1))
+  }
+})
+
+# Create a new dataframe, q_scores_df, by combining sampled_q_maternal and sampled_q_paternal
+q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
+
+# Plot a scatterplot of sampled_q_paternal versus sampled_q_maternal to visualize the relationship between the two variables where they are paired based on extreme q score preference
+plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
+
+# To plot offspring q scores, take the q scores of the parents from the previous data frame and average them 
+q_scores_df$averaged_q_scores <- (q_scores_df$sampled_q_maternal + q_scores_df$sampled_q_paternal) / 2
 
@@
 
-# Perform Spearman rank correlation test
+# Perform Spearman rank correlation test between maternal and paternal q scores
 cor_test_result <- cor.test(q_scores_df$sampled_q_maternal, 
@@
 ```
-
-### Assortative Mating - Extreme Phenotype Preference
-```{r Assortative Mating - Extreme Phenotype Preference}
-# Set a seed for reproducibility
-set.seed(123)
-
-# Create two datasets, `q_maternal` and `q_paternal`, each with 1000 data points following a beta distribution
-q_maternal <- rbeta(1000, 0.3, 0.3)
-q_paternal <- rbeta(1000, 0.3, 0.3)
-
-# Display simulated dataset distribution before sampling
-hist(q_paternal)
-hist(q_maternal)
-
-# Create a new dataset, sampled_q_maternal, by sampling 500 q scores without replacement from the q_maternal dataset
-sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
-
-# Create a new variable, sampled_q_paternal, by sampling 500 q scores from the q_paternal dataset
-# Use the ifelse() function to set the probability of selecting a q score from q_paternal based on the corresponding q score in sampled_q_maternal
-sampled_q_paternal <- sapply(sampled_q_maternal, function(x) {
-  if (x <= 0.5) {
-    # If the maternal q score is less than or equal to 0.5, sample from the lower half of the paternal q scores with higher probability
-    sample(q_paternal[q_paternal <= 0.5], 1, prob = ifelse(q_paternal[q_paternal <= 0.5] <= 0.25, 0.9, 0.1))
-  } else {
-    # If the maternal q score is greater than 0.5, sample from the upper half of the paternal q scores with higher probability
-    sample(q_paternal[q_paternal > 0.5], 1, prob = ifelse(q_paternal[q_paternal > 0.5] >= 0.75, 0.9, 0.1))
-  }
-})
-
-# Create a new dataframe, q_scores_df, by combining sampled_q_maternal and sampled_q_paternal
-q_scores_df <- data.frame(sampled_q_maternal, sampled_q_paternal)
-
-# Plot a scatterplot of sampled_q_paternal versus sampled_q_maternal to visualize the relationship between the two variables where they are paired based on extreme q score preference
-plot(sampled_q_maternal, sampled_q_paternal, xlab="Sampled Maternal q scores", ylab="Sampled Paternal q scores", main="Scatterplot of Maternal vs Paternal q scores")
-
-q_scores_df$averaged_q_scores <- (q_scores_df$sampled_q_maternal + q_scores_df$sampled_q_paternal) / 2
-# Estimate the density of averaged_q_scores
-# density_averaged_q_scores <- density(q_scores_df$averaged_q_scores)
-
-# Plot the density plot which shows offspring q scores based on extreme phenotype assortative mating
-plot(density(q_scores_df$averaged_q_scores), 
-     main = "Density Plot of Offspring q Scores",
-     xlab = "q score",
-     ylab = "Density")
-
-# Add a rug plot to the bottom of the density plot to show the individual data points that make up the distribution
-rug(q_scores_df$averaged_q_scores)
-
-# Perform Spearman rank correlation test
-cor_test_result <- cor.test(q_scores_df$sampled_q_maternal, 
-                           q_scores_df$sampled_q_paternal, 
-                           method = "spearman")
-
-# Print the correlation test result
-print(cor_test_result)
-```

### 2024-11-30 Registered Report.Rmd
Improved formatting and added more annotations.

@@
 
-### Description
+## Proposed Analysis Pipeline:
 
-Code looks to simulate assortative mating between two species to potentially form hybrid offspring.
+### Preprocessing Steps:
+
+There will be separate simulations for two different cases of assortative mating.
+In both cases, the initial maternal and paternal q scores will be simulated following a beta distribution.
+A beta distribution is ideal for q scores as they can take on values between 0 and 1.
+The initial pattern for the beta distribution of q scores will be a bimodal pattern with peaks around 0 and 1 indicating that a majority of the individuals are parental species rather than hybrids.
+1000 maternal and paternal q scores will be simulated for a total of 2000 which is consistent with literature on the topic of assortative mating simulation.
+Using the sample function in R, maternal q scores will be sampled from the initial maternal q score dataset before being matched with sampled paternal q scores.
+In the strict case of assortative mating, maternal q scores will be paired with paternal q scores of a similar value.
+In the case of mating based on extreme phenotype, the maternal q scores will be paired with paternal q scores with extreme values (e.g.
+maternal q scores less than 0.5 will be paired with paternal q scores closer to 0 and maternal q scores higher than 0.5 will be paired with paternal q scores closer to 1).
+Using these pairings for both assortative mating types, the parental q scores will be averaged to generate offspring q scores.
+No statistical power analysis is proposed as the thesis involves the simulation of data rather than sampling associated with laboratory and fieldwork methodology.
+
+It is important to note that the aforementioned steps are done to approximate a dataset that would otherwise be simulated using forward genetic simulation software.
+This requires simplifying assumptions such as blended inheritance for offspring, matching based on q scores, and the exclusion of phenotypes.
+
+### Planned Analysis:
+
+#### Spearman’s rank correlation test:
+
+This test assesses the correlation between the paired maternal and paternal q scores.
+This non-parametric correlation test will be performed instead of Pearson’s correlation test as q scores are not normally distributed.
+In both cases of assortative mating, a correlation is expected, with the null hypothesis being random mating.
+A positive correlation would indicate a preference for mates with a similar q score.
+
+#### Beta regression:
+
+A beta regression will be used as the response variable (q score) follows a beta distribution.
+Potential parameters to explore include hybridization rates, strength of preference, and the influence of mating pool availability on preference.
+
+It should be noted that due to the simplifying assumptions made in this code, a beta regression was not performed.
+Furthermore, strength of preference was preset for both cases of assortative mating based on sampling and matching probabilities for q scores.
+
+### Code Description:
+
+The code simulates assortative mating between two species to potentially form hybrid offspring.
 This assortative mating will be based on q scores which can range from 0 (which indicates species 1) to 1 (which indicates species 2) and intermediate values which indicate hybrids of the two.
@@
 The code will simulate individual maternal and paternal datasets and then pair them up based on either of the two assortative mating definitions previously mentioned.
-The aformentioned pairing will be visualized in a scatterplot to show the mating patterns.
+The aforementioned pairing will be visualized in a scatter plot to show the mating patterns.
 Then, offspring will be generated based on those previous mating patterns that will be a blend of the q scores of both parents.
@@
 # Create a new dataset, sampled_q_maternal, by sampling 500 q scores without replacement from the q_maternal dataset
-sampled_q_maternal <- sample(q_maternal, 500, replace=FALSE)
+sampled_q_maternal <- sample(q_maternal, 500, replace = FALSE)
 

### 2024-12-14 Paper_Summary.Rmd
Paper Summary Final.

@@
 Two different environments were set up: a small vial and a larger Ziploc container, both containing food.
-Each environment had ten replicates, with 37 male and 37 female flies in each.
+Each environment had ten replicates, with 34 male and 34 female flies in each.
 The female flies were divided into two categories, small and large, based on manipulated egg density.
@@
 Data from each arena were then aggregated for analysis.
+Sexual activity behaviour, and feeding observations were the response variables whereas the simple and complex environments were the predictor variables for this assay.
+A two-sample t-test was done to compare the difference in sexual activity score of the two female types between the two mating environments.
+Furthermore, a general linear model was used to analyze how feeding behaviour was influenced by mating arena, female size, and their interaction.
 
@@
 Total adult fitness, encompassing both fecundity and survival, was calculated by multiplying these two values.
+The response variable was total adult fitness and the predictor variables were the mating arena and male exposure type.
+A general linear model was used to determine the variation in total adult fitness with the fixed effects being mating arena, male exposure, and their interaction.
 
@@
 in cages.
-It made sense to use the two-sample t-test to compare means between the two environments as one of the goals of this paper was to compare how the environment changed the sexual interactions of Drosophila melanogaster.
+It made sense to use the two-sample t-test to compare means between the two environments, as one of the goals of this paper was to compare how the environment changed the sexual interactions of Drosophila melanogaster.
 It is also important to note that the authors took steps to avoid violating one of the assumptions of the two-sample t-test (independence) by creating a single sexual activity score for each replicate (for both small and large females) as the observations of average number of sexual encounters within the cages and vials were not independent.
+A general linear model was also used to analyze the influence of mating arena, female size, and their interaction on feeding observations.
+What the authors observed was that large females fed more often in cages than in vials whereas small females did not.
+This indicated that there was a significant interaction between mating arena and female size (F1,36 = 47.36, p < 0.001).
+The authors also found that feeding behaviours were higher in cages than vials when pooling across females (F1,36 = 21.30, p < 0.001).
+It made sense for the authors to use a general linear model as there were multiple independent variables and a single dependent variable (feeding).
+They also wanted to see how the independent variables interacted with each other to influence the feeding behaviours and they found a significant interaction between mating arena and female size.
+Similarly to the previous t-test, the goal of the general linear model was to determine the influence of the environment on female fruit flies when exposed to males.
 
@@
 
-Insert sample text.
+The statistical method used for assay 2 was also a general linear model.
+This time the authors looked at the influence of mating arena, male exposure, and their interaction on total adult fitness of female fruit flies.
+What the authors found was that under lower male exposure, female fitness increased (F1,134 = 22.70, p < 0.0001).
+They also found that females in the complex cage arena had higher fitness than in vials (F1,134 = 110.43, p < 0.0001).
+Finally, the authors also concluded that there was a significant interaction between male exposure and mating arena suggesting that the impact of male exposure is higher in vials (F1,134 = 7.49, p < 0.007).
+The usage of the general linear model was a prudent one as it provided key information about male harm and how it is mediated by the environment which was a goal of this study.
+As there were multiple independent variables that could interact with each other to influence the predictor variable, it made sense to use a general linear model as the statistical methods.
 
@@
 
-For the methods and results in Assay 1, the dataset was simulated and analysized in two different ways corresponding to Figures 1a and 1b in the original paper.
-The corresponding code is seperated into two blocks, one for Figure 1a and the other for Figure 1b.
+For the methods and results in Assay 1, the dataset was simulated and analysed corresponding to Figure 1a in the original paper.
+The corresponding code is separated into two blocks, one for Figure 1a simulation and the other for plotting.
+Note: Figure 1b is not shown as it had similar steps to Figure 2 which was also plotted and analyzed.
 
@@
 
-# Set the average differences in activity scores
+# Set the average differences in activity scores (this step was done in the paper without showing their work so this was an estimate of what they did)
 mean_diff_simple <- 6
@@
 
-# Generate data for each vial/cage (For loops are harder in R than in Python in my opinion!)
+# Generate data for each vial/cage, repeating until i = N of vials = which should be 10 (For loops are harder in R than in Python in my opinion!)
 for (i in 1:n_vials) {
@@
     vial_cage_id = i,
-    activity_score = rnorm(1, mean = 10 + mean_diff_simple/2, sd = sd_diff)
+    activity_score = rnorm(1, mean = 10 + mean_diff_simple/2, sd = sd_diff) # 
   ))
@@
 
-# Perform the two-sample t-test 
+# Perform the two-sample t-test looking for difference in activity score between vial and cage groups
+# Note: t-test should find a significant difference in means between the groups to match the value found in the paper. Approx t = 4
 t.test(score_difference ~ arena_type, data = activity_data)
@@
 ```{r Assay 1: Figure 1a. Plot}
-# Trait of Interest:
-
-# Sample Data (replace with your actual data)
+# Code that generates a plot similar to figure 1a of the paper
+# Data that is broken up into vial and cage with values for small females and large females. 
 data <- data.frame(
@@
   Size = factor(rep(c('Small', 'Large'), 2), levels = c('Small', 'Large')),
-  Mean = c(4, 10, 6, 5),
-  SD = c(1, 1, 1, 1)  # Standard deviation (replace with your actual values)
+  Mean = c(4, 10, 6, 5), # mean value of sexual interactions as found in the literature
+  SD = c(1, 1, 1, 1)    # Standard deviation of 1 as a default as none was indicated in the paper
 )
 
+# Arena replicates of 10 of each mating arena = N of 20 to calculate standard error
+n <- 20  
+data$SE <- data$SD / sqrt(n)  # Calculate Standard Error
+
 # Create the bar graph with error bars (Environment on x-axis)
-ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0, 2, 4, 6, 8, 10)) +
+ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + 
+  scale_y_continuous(breaks=c(0, 2, 4, 6, 8, 10)) +
   geom_bar(stat = 'identity', position = 'dodge') +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), 
+  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),  # Use SE here
                 width = 0.2, position = position_dodge(0.9)) +
@@
 
-Sample Text. Please fill in.
-
-#### Figure 1b: Data Simulation and Analysis
-
-```{r Assay 1: Figure 1b. Data Simulation & Analysis}
-# RNG time! - Let's keep it consistent 
-set.seed(123)
-
-# Number of observations made - 2 females, 3 times a day for 3 days * ten replicates for each arena = 180 observations per environment type.
-N = 180
-
-# Changing the effect sizes to get the same values as the paper
-effect_arena <- 2.0       # Main effect of Arena
-effect_size <- 0.5        # Main effect of Size
-effect_interaction <- 2.0  # Interaction effect
-
-# Data simulation time! - Different means for different groups
-feeding_count <- c(
-  rnorm(N, mean = 6.5, sd = 1), # Large females in vials
-  rnorm(N, mean = 13.5 + effect_arena, sd = 1), # Large females in cages
-  rnorm(N, mean = 10 + effect_size, sd = 1), # Small females in vials
-  rnorm(N, mean = 8.5 + effect_arena + (effect_size * (1 + effect_interaction)), sd = 1) # Small females in cages
-)
-
-# Making sure that FeedingCount is a positive float (can't eat negative amounts!)
-feeding_count <- ifelse(feeding_count <= 0, feeding_count + 0.1, feeding_count)
-
-# Let's make all of this into a data frame - matrix time :(
-data <- data.frame(
-  Arena = factor(rep(c('vial', 'cage'), each = N * 2)),
-  Size = factor(rep(c('large', 'small'), times = N * 2)),
-  FeedingCount = feeding_count
-)
-
-model_gaussian <- glm(FeedingCount ~ Arena * Size, data = data, family = gaussian(link = 'log'))
-
-summary(model_gaussian)
-
-# Use lmerTest to get F-statistics and p-values
-anova_result <- anova(model_gaussian, test = "F")
-
-# Print the results
-print(anova_result)
-
-```
-
-#### Figure 1b: Plot of Results
-
-```{r Assay 1: Figure 1b. Plot}
-
-# Spawning and simulating useful variables to create the plot for Figure 1b
-MeanFeeding_Small<-c(10, 9)
-MeanFeeding_Large<-c(6.5, 13.5)
-EnvironmentType<-c('Vial', 'Cage')
-x <- c(1, 2)
-SD = c(1, 1, 1, 1)
-
-# Creating a plot that is similar to Figure 1b. in the paper
-plot(MeanFeeding_Small, axes = FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim = c(4, 16), pch = 15, col = '#B6D0E2', type = 'o', lwd = 2, cex = 1.5)
-  axis(2)
-  axis(1, at = seq_along(MeanFeeding_Small), labels = as.character(EnvironmentType), las = 0)
-  lines(c(1, 2), MeanFeeding_Large, col='#F1E3D3', pch = 17, type = 'o', lty = 2, lwd = 2, cex = 1.5)
-  title(main = '(b)')
-
-  # Vertical error bars the good ole fashion way
-  arrows(x0 = x, y0 = MeanFeeding_Small - SD, x1 = x, y1 = MeanFeeding_Small + SD, code = 3, angle = 90, 
-         length = 0.1)
-  arrows(x0 = x, y0 = MeanFeeding_Large - SD, x1 = x, y1 = MeanFeeding_Large + SD, code = 3, angle = 90, 
-         length = 0.1)
-
-  # Creating a legend that labels the female size types with different colours - corresponds to the colours in      the previous graph
-  legend('topleft', inset = 0.05, c('Small', 'Large'), lty = c(1, 2), col =c ('#B6D0E2','#F1E3D3'), 
-         lwd = 2, title = "Female Size")
-
-  # Turn plot outline into a box to make it more cutesy
-  box()
-```
-
-#### Interpretation of Results in Figure 1b
-
-Sample text. Please fill in.
+The results indicated that the means (difference in sexual activity) of the vial group and the cage group were significantly different as indicated by the values (t18 = 5.89, p < 0.001).
+To the authors this suggested that the difference in male sexual attention towards large and small females was different in vials vs.
+in cages.
+Most of the increase in sexual attention was directed towards large females in vials.
 
 #### Assay 2
+
+Simulating and analyzing the data from Assay 2 which corresponds to figure 2 of the paper.
 
@@
 ```{r Assay 2: Figure 2. Data Simulation & Analysis}
-# Simulated data
+# Simulating the data from Assay 2 of the paper that makes up figure 2. The analysis is the general linear model
+# Simulated data following a normal distribution with an N of 35 matching that found in the paper. Mean is 1 so that it can be adjusted later to match the results of the paper. 
 low.vial <- rnorm(35, 1, 0.1)
@@
 response <- c(
-  24 * low.vial + rnorm(35, mean = 0, sd = 0.1),  # Low Vial
-  17 * high.vial + rnorm(35, mean = 0, sd = 0.1), # High Vial
-  30 * low.cage + rnorm(35, mean = 0, sd = 0.1),  # Low Cage
-  28 * high.cage + rnorm(35, mean = 0, sd = 0.1)   # High Cage
+  24 * low.vial + rnorm(35, mean = 0, sd = 0.1),  # Low Vial with adjusted mean to match that found in paper
+  17 * high.vial + rnorm(35, mean = 0, sd = 0.1), # High Vial with adjusted mean to match that found in paper
+  30 * low.cage + rnorm(35, mean = 0, sd = 0.1),  # Low Cage with adjusted mean to match that found in paper
+  28 * high.cage + rnorm(35, mean = 0, sd = 0.1)   # High Cage with adjusted mean to match that found in paper
 )
@@
 
-# Fit the two-way ANOVA model
+# Fit the two-way ANOVA model which shows similar results to the paper with lower fitness with higher male exposure and lower fitness in the vial groups
 anova_model <- aov(response ~ group + type + group:type, data = data)
@@
 
-Insert plot here.
+``` {r Assay 2: Figure 2. Plot}
+# Simulated data following a normal distribution with an N of 35 matching that found in the paper. Mean is 1 so that it can be adjusted later to match the results of the paper. 
+low.vial <- rnorm(35, 1, 0.1)
+high.vial <- rnorm(35, 1, 0.1)
+low.cage <- rnorm(35, 1, 0.1)
+high.cage <- rnorm(35, 1, 0.1)
+
+# Combine into a single response variable
+response <- c(
+  24 * low.vial + rnorm(35, mean = 0, sd = 0.1),  # Low Vial with adjusted mean to match that found in paper
+  17 * high.vial + rnorm(35, mean = 0, sd = 0.1), # High Vial with adjusted mean to match that found in paper
+  30 * low.cage + rnorm(35, mean = 0, sd = 0.1),  # Low Cage with adjusted mean to match that found in paper
+  28 * high.cage + rnorm(35, mean = 0, sd = 0.1)    # High Cage with adjusted mean to match that found in paper
+)
+
+# Divide data points in groups (low vs high) and type (vial vs cage) so that they can be ordered on the plot later
+data <- data.frame(
+  response = response,
+  group = factor(rep(c("Low", "High", "Low", "High"), each = 35)),
+  type = factor(rep(c("Vial", "Vial", "Cage", "Cage"), each = 35), levels = c("Vial", "Cage")) # Specify the order of levels
+)
+
+# Fit the two-way ANOVA model
+anova_model <- aov(response ~ group + type + group:type, data = data)
+
+# Summary of the ANOVA
+summary(anova_model)
+
+# Calculate means and standard errors to be plotted
+data_summary <- aggregate(response ~ group + type, data = data, FUN = function(x) c(mean = mean(x), se = sd(x)/sqrt(length(x))))
+data_summary <- do.call(data.frame, data_summary)
+
+# Create the interaction plot with error bars matching that found in figure 2 of the paper
+interaction.plot(data_summary$type, data_summary$group, data_summary$response.mean, 
+                 type = "b", # "b" for both lines and points
+                 pch = c(1,16), # Different point symbols
+                 col = c("red","blue"), # Different colors for lines and points
+                 xlab = "Mating Arena (Vial/Cage)", 
+                 ylab = "'Total' Adult Fitness",
+                 trace.label = "Male Exposure",
+                 main = "Interaction of Mating Arena on Adult Fitness",
+                 ylim = c(min(data_summary$response.mean - 2*data_summary$response.se), max(data_summary$response.mean + 2*data_summary$response.se))) # Adjust ylim to accommodate error bars
+
+# Error bars that match the standard error of 1 found in the paper's error bars. 
+arrows(x0 = as.numeric(data_summary$type), # Create error bars using the arrow function in r
+       y0 = data_summary$response.mean - data_summary$response.se, # minus 1 se
+       x1 = as.numeric(data_summary$type), 
+       y1 = data_summary$response.mean + data_summary$response.se, # plus 1 se
+       length = 0.05, angle = 90, code = 3)
+
+```
 
